---
title: "lolo"
author: "Christian M"
date: "October 2, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

**Reading in the data**

1. Make sure tidyverse is loaded (to use the pipe function %>%)
2. List files in directory (within data folder)
3. Combine each file into one data frame (or tibble)

```{r}
lolo_data <- fs::dir_ls("data/lolo")
lolo_all_raw <- lolo_data %>% map_dfr(read_csv)
```

**Decluttering**

1. Omitting data (SET_CN, PLOT_CN, TREE_CN, TAG_ID, AZIMUTH, DISTANCE, SNAD_DECA and disturbance items.

2. There are `r length(unique(lolo_omit$SETTING_ID))` stands in the lolo dataset.

3. There are `r length(unique(lolo_omit$MEASUREMENT_DATE))` different measurement dates.
a. Goal - Separate out dates by year.
  Options: lubridate package?


```{r}
lolo_omit <- select(lolo_all_raw, SETTING_ID, MEASUREMENT_DATE, MEASUREMENT_NO, PURPOSE_CODE, PLOT, SUBPLOT, LIVE_DEAD, SPECIES_SYMBOL, DIAMETER, TPA_EQUIV, BASAL_AREA_EQUIV, HEIGHT, CROWN_RATIO, CROWN_CLASS, AGE, RADIAL_GROWTH, HEIGHT_GROWTH)

unique(lolo_omit$SETTING_ID)
length(unique(lolo_omit$SETTING_ID))
```

**Separating Some Variable - Placeholder**

```{r}

```

