---
title: "lolo"
author: "Christian M"
date: "October 2, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

**Reading in the data**

1. Make sure tidyverse is loaded (to use the pipe function %>%)
2. List files in directory (within data folder)
3. Combine each file into one data frame (or tibble)

```{r}
lolo_data <- fs::dir_ls("data/lolo")
lolo_all_raw <- lolo_data %>% map_dfr(read_csv)
```

**Decluttering**



```{r}
#omitting data that isn't immediately useful

lolo_omit <- select(lolo_all_raw, SETTING_ID, MEASUREMENT_DATE, MEASUREMENT_NO, PURPOSE_CODE, PLOT, SUBPLOT, LIVE_DEAD, SPECIES_SYMBOL, DIAMETER, TPA_EQUIV, BASAL_AREA_EQUIV, HEIGHT, CROWN_RATIO, CROWN_CLASS, AGE, RADIAL_GROWTH, HEIGHT_GROWTH)

#cleaning up dates, adding column with proper date format.

lolo1 <- mutate(lolo_omit, mdate = as.Date(MEASUREMENT_DATE, format = "%m/%d/%Y"))

#adding column with measure year

lolo1a <- mutate(lolo1, myear = format(mdate, format = "%Y"))

unique(lolo_omit$SETTING_ID)
length(unique(lolo_omit$SETTING_ID))
```

1. Omitting data (SET_CN, PLOT_CN, TREE_CN, TAG_ID, AZIMUTH, DISTANCE, SNAD_DECA and disturbance items.

2. There are `r length(unique(lolo_omit$SETTING_ID))` stands in the lolo dataset.

3. There are `r length(unique(lolo_omit$MEASUREMENT_DATE))` different measurement dates.
a. clean up dates, extract useful info




```{r}
stand1 <- lolo1a[lolo1a$SETTING_ID == "01160304010002", ]
stand1 <- stand1[stand1$myear == 1987, ]
stand1 <- stand1[round(stand1$PLOT, -2) == 100, ]
```


```{r}
ggplot(stand1, aes(x = DIAMETER, y = HEIGHT, color = SPECIES_SYMBOL)) + geom_point() + geom_smooth(se = F)
```



```{r}

#Creating Height vs. Diameter plot for selected stand, year, cluster.

hdplot <- function(settingID, year, cluster) {

stand1 <- lolo1a[lolo1a$SETTING_ID == settingID, ]
stand1 <- stand1[stand1$myear == year, ]
stand1 <- stand1[round(stand1$PLOT, -2) == cluster, ]

ggplot(stand1, aes(x = DIAMETER, y = HEIGHT, color = SPECIES_SYMBOL)) + geom_point() + geom_smooth(se = F)}

hdplot("01160304010002", 1987, 200)
```



```{r, fig.show='hold', out.width='50%'}

#Reproducing the above with dplyr

hdplot1 <- function(settingID, year, cluster) {
  
a <- filter(lolo1a, SETTING_ID == settingID,  myear == year, round(PLOT, -2) == cluster)

ggplot(a, aes(x = DIAMETER, y = HEIGHT, color = SPECIES_SYMBOL)) + geom_point() +
  geom_smooth(se = F)
}
hdplot1("01160304010002", 1987, 200)
hdplot1("01160304010002", 2000, 200)
```

```{r}
# Diameter vs. year boxplots for a given cluster. It would make sense to show mean diameter or QMD and exclude the sublplot data


a <- lolo1a %>% filter(SETTING_ID == "01160304010002", round(PLOT, -2) == 100, SPECIES_SYMBOL == "ABLA")

ggplot(a, aes(x = myear, y = DIAMETER)) + geom_boxplot()

```




```{r}
#This is incomplete and needs work. Goal was to find a L/D ratio and then plot it. 


#data set then filter cluster then count unique values in livedead then extract the vector then extract desired value from vecture (Live) / "" (Dead)
lolo1a %>% filter(round(PLOT, -2) == 100) %>% count(LIVE_DEAD) %>% pull(n) %>% nth(2)/
lolo1a %>% filter(round(PLOT, -2) == 100) %>% count(LIVE_DEAD) %>% pull(n) %>% nth(1)

ldratioplot <- function(cluster) {}
  
ratio <- lolo1a %>% filter(SETTING_ID == "01160304010002", round(PLOT, -2) == 200) %>% mutate(live = lolo1a %>% count(LIVE_DEAD) %>% pull(n) %>% nth(2), dead = lolo1a %>% count(LIVE_DEAD) %>% pull(n) %>% nth(1))

ggplot(ratio, aes(x = myear, y = (live/dead))) + geom_point()


```

```{r}
#For any cluster year, how many live trees/acre??

#pull data, then select rows with tpa (expansion factor) = 20, then select desired variables (columns) then, group by (this is only a formality), then turn L's into 1's while omitting D's, then count the 1's and compress each plot into one row per measure year, then rename variable

#lolo1a %>% filter(TPA_EQUIV == 20) %>% select(SETTING_ID, PLOT, myear, LIVE_DEAD, TPA_EQUIV) %>% group_by(SETTING_ID, cluster = PLOT, myear, LIVE_DEAD, tpa = TPA_EQUIV) %>% summarize(LIVE_DEAD = str_count(LIVE_DEAD, "L")) %>% summarize(LIVE_DEAD = sum(LIVE_DEAD)) %>% rename(live_trees = LIVE_DEAD) %>% mutate(live_tpa = live_trees*tpa)

#live_trees <- function(stand, cluster){
  
#a <- 
lolo1a %>% filter(SETTING_ID == "01160304010002", TPA_EQUIV == 20) %>% select(SETTING_ID, PLOT, myear, LIVE_DEAD, TPA_EQUIV) %>% group_by(SETTING_ID, cluster = PLOT, myear, LIVE_DEAD, tpa = TPA_EQUIV) %>% summarize(LIVE_DEAD = str_count(LIVE_DEAD, "L")) %>% summarize(LIVE_DEAD = sum(LIVE_DEAD)) %>% rename(live_trees = LIVE_DEAD) %>% mutate(live_tpa = live_trees*tpa)

ggplot(a, aes(x = myear, y = live_tpa, color = cluster)) + geom_point()
##ggplot(a, aes(x = myear, y = LIVE_DEAD, color = SETTING_ID)) + geom_point()
#}

#The above presents a tibble which shows the count of live trees in a plot cluster at a given year. Current issue: the average tpa column is confounded by subplot data

#lolo1a %>% select(SETTING_ID, PLOT, myear, LIVE_DEAD, TPA_EQUIV) %>% group_by(SETTING_ID, PLOT, myear, LIVE_DEAD, TPA_EQUIV) %>% summarize(LIVE_DEAD = str_count(LIVE_DEAD, "L")) %>% summarize(LIVE_DEAD = sum(LIVE_DEAD))

#%>% summarize(cluster = round(cluster, -2))
```

```{r}
# read the data and get some new variables
source("1.readformat.R")
head(lolo1)

# drop the SE and CI purpose code data - these are done in "off" years  
#  and seem to involve variable radius plot sampling; I think they
#  represent other data coll efforts --> need to check tho
lolo1 <- lolo1[!(lolo1$PURPOSE_CODE %in% c("SE","CI")),]
table(lolo1$PURPOSE_CODE) # need to figure out what "RR" is? Replacement?

# check whether all installations had 3 plots per cluster; info is needed because 
#  TPA is recorded as 20 tr/ac not 20/3 tr/ac
plot_structure <- aggregate(lolo1$PLOT,
                               by=list(SETTING_ID=lolo1$SETTING_ID,
                                       PURPOSE_CODE=lolo1$PURPOSE_CODE,
                                       myear=lolo1$myear,
                                       cluster=lolo1$cluster),
                            FUN=function(x){length(unique(x))})
unique(plot_structure$x)


# check whether all plots had 3 subplots
lolo1a <- lolo1[!is.na(lolo1$SUBPLOT),]
subplot_structure <- aggregate(lolo1a$SUBPLOT,
                               by=list(SETTING_ID=lolo1a$SETTING_ID,
                                       PURPOSE_CODE=lolo1a$PURPOSE_CODE,
                                       cluster=lolo1a$cluster,
                                       PLOT=lolo1a$PLOT),
                               FUN=function(x){length(unique(x))})
unique(subplot_structure$x)


# just live trees
lolo1_live <- lolo1[!is.na(lolo1$TPA_EQUIV) & lolo1$LIVE_DEAD=="L",]
unique(lolo1_live$TPA_EQUIV)

# what is RR? Looks like it used 20th acre plots for large and 100th acre for small?
unique(lolo1_live$TPA_EQUIV[lolo1_live$PURPOSE_CODE=="RR"])

# all other small tree plots appear to 300th acre
unique(lolo1_live$TPA_EQUIV[lolo1_live$PURPOSE_CODE!="RR" & !is.na(lolo1_live$SUBPLOT)]) %% 300
unique(lolo1_live$TPA_EQUIV[lolo1_live$PURPOSE_CODE!="RR" & is.na(lolo1_live$SUBPLOT)]) 

# fix tpa to account for multiplicity of plots, subplots per cluster
lolo1_live$TPA_EQUIV <- with(lolo1_live, TPA_EQUIV/ifelse(is.na(SUBPLOT),3,9))

# obtain basal area equivalents
lolo1_live$BA_EQUIV <- with(lolo1_live,
                            ifelse(!is.na(DIAMETER),DIAMETER,0)^2 * 
                              (pi/576)* TPA_EQUIV )

# sum to the cluster level
stand_char <- aggregate(lolo1_live[,c("TPA_EQUIV","BA_EQUIV")],
                        by=list(SETTING_ID=lolo1_live$SETTING_ID,
                                myear=lolo1_live$myear,
                                PURPOSE_CODE=lolo1_live$PURPOSE_CODE,
                                cluster=lolo1_live$cluster),
                        sum)
stand_char$DBH_Q <- with(stand_char, sqrt(576*BA_EQUIV/TPA_EQUIV/pi) )

```